---
title: "Regression Project Code"
author: "Jill Reiner"
date: "6/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(forcats)
library(ggcorrplot)
```

```{r}
cfb <- read.csv("http://www.stat.cmu.edu/cmsac/sure/materials/data/regression_projects/cfb_2019_games.csv")

cfb[is.na(cfb$excitement_index),]

cfb <- cfb[-c(268,625),]

cfb$home_1sthalf <- cfb$home_1_pts + cfb$home_2_pts 
cfb$home_2ndhalf <- cfb$home_3_pts + cfb$home_4_pts

cfb$away_1sthalf <- cfb$away_1_pts + cfb$away_2_pts
cfb$away_2ndhalf <- cfb$away_3_pts + cfb$away_4_pts

cfb$margin_1sthalf <- cfb$home_1sthalf - cfb$away_1sthalf
cfb$margin_game <- cfb$home_points - cfb$away_points

cfb$comeback <- (cfb$margin_game / cfb$margin_1sthalf) < 0

cfb$overtime <- !((cfb$home_1_pts + cfb$home_2_pts + cfb$home_3_pts + cfb$home_4_pts) == cfb$home_points |
                    (cfb$away_1_pts + cfb$away_2_pts + cfb$away_3_pts + cfb$away_4_pts) == cfb$away_points)

```

```{r}
cfb %>%
  ggplot(aes(x = excitement_index)) +
  geom_histogram(binwidth = .9) +
  theme_bw()
```

```{r}
cfb %>%
  group_by(conference_game) %>%
  summarize(avg_ei = mean(excitement_index)) %>%
  ggplot(aes(x = conference_game, y = avg_ei)) +
  geom_bar(stat = "identity") +
  theme_bw()
```

```{r}
cfb %>%
  mutate(totalpoints = away_4_pts + home_4_pts + away_3_pts + home_3_pts + away_2_pts + home_2_pts + away_1_pts + home_1_pts) %>%
  ggplot(aes(x = excitement_index, y = totalpoints)) +
  geom_point() +
  stat_smooth() +
  theme_bw()
```

Not very good. Total points could either be a complete blowout (not very exciting) or a close high scoring game (exciting).

```{r}
cfb_conference <- cfb %>%
  filter(home_conference %in% c("ACC", "Big Ten", "Big 12", "Pac-12", "SEC")) %>%
  group_by(home_conference) %>%
  summarize(avg_ei = mean(excitement_index)) %>%
  ggplot(aes(x = home_conference)) +
  geom_bar(aes(y = avg_ei),
           stat = "identity") +
  theme_bw() +
  labs(x = "Conference", y = "Average Excitement Index per Game")
  cfb_conference
```

```{r}
mean(cfb$excitement_index)
```

```{r}
cfb_conference3 <- cfb %>%
  filter(home_conference %in% c("ACC", "Big Ten", "Big 12", "Pac-12", "SEC") |
           away_conference %in% c("ACC", "Big Ten", "Big 12", "Pac-12", "SEC")) %>%
  filter(conference_game == TRUE)
summary(cfb_conference3)
```

```{r}
cfb_con <- cfb_conference3 %>%
  ggplot(aes(x = excitement_index,
             y = home_conference)) +
  geom_density_ridges(jittered_points = TRUE, position = position_points_jitter(width = 0.05, height = 0), point_shape = '|', point_size = 3, point_alpha = 1, alpha = 0.7,
                      quantile_lines = TRUE,
                      quantile_fun=function(x,...)mean(x)) +
  geom_vline(xintercept = mean(cfb_conference3$excitement_index), color = "blue") +
  labs(x = "Excitement Index", y = "Conference",
       title = "Pac-12, Big 12, and the ACC had the most exciting games in 2019") +
  theme_bw() 
cfb_con
ggsave("con_dist.jpeg", width = 8, height = 6)
```

```{r}
lm1 <- lm(data = cfb, excitement_index ~ score_difference + comeback * margin_1sthalf + conference_game)
summary(lm1)
```

```{r}
cfb_model_data_ourvars <- cfb %>%
  dplyr::select(
                score_difference,
                home_1sthalf,
                margin_1sthalf,
                comeback,
                overtime,
                excitement_index)
cfb_cor_matrix_ourvars <- cor(cfb_model_data_ourvars)

ggcorrplot(cfb_cor_matrix_ourvars, 
           type = "lower",
           lab = TRUE)
ggsave("corr_plot.jpeg", width = 7, height = 6)
```

```{r pretty-cor, echo = TRUE, eval = FALSE}
round_cor_matrix <- 
  round(cor(cfb_model_data_ourvars), 2) #<<
ggcorrplot(round_cor_matrix, 
           hc.order = TRUE,#<<
           type = "lower",#<<
           lab = TRUE)#<<
```

```{r select-preds}
#library(ggdendro)
#cfb_ex_vars <- dplyr::select(cfb_model_data, -excitement_index)
#exp_cor_matrix <- cor(cfb_ex_vars)

#cor_dist_matrix <- 1 - abs(exp_cor_matrix)
#cor_dist_matrix <- as.dist(cor_dist_matrix)

#cfb_exp_hc <- hclust(cor_dist_matrix, #<<
                     #"complete") 
#ggdendrogram(cfb_exp_hc, #<<
             #rotate = TRUE, #<<
             #size = 2)#<<
```

```{r}
#library(dendextend)
#cor_dist_matrix %>%
 #hclust() %>%
  #as.dendrogram() %>%
  #set("branches_k_col", 
      #k = 2) %>% 
  #set("labels_cex", .9) %>%
  #ggplot(horiz = TRUE)
```

Home and away 4th Q points in their own cluster? Interesting development.

```{r pairsplot, eval = FALSE}
library(GGally) #<<
ggpairs(cfb, #<<
        columns =
          c("excitement_index", "score_difference",
            "margin_1sthalf", "comeback",
            "margin_game", "overtime")) #<<
```


```{r init-folds}
set.seed(2020)
cfb <- cfb %>%
  mutate(test_fold = sample(rep(1:5, length.out = n())))
```

```{r}
get_cv_preds <- function(model_formula, data = cfb) {
  # generate holdout predictions for every row based season
  map_dfr(unique(data$test_fold), 
          function(holdout) {
            # Separate test and training data:
            test_data <- data %>%
              filter(test_fold == holdout)
            train_data <- data %>%
              filter(test_fold != holdout)
            
            # Train model:
            reg_model <- lm(as.formula(model_formula), data = train_data)
            
            # Return tibble of holdout results:
            tibble(test_preds = predict(reg_model, newdata = test_data),
                   test_actual = test_data$excitement_index,
                   test_fold = holdout) 
          })
}

```

```{r}
ourvars_minusmargingames_cv_preds <- get_cv_preds("excitement_index ~ score_difference + margin_1sthalf + comeback")
ourvars_cv_preds <- get_cv_preds("excitement_index ~ score_difference * margin_1sthalf + comeback + overtime")
all_int_cv_preds <- get_cv_preds("excitement_index ~ score_difference * margin_1sthalf + comeback")
margin_only_cv_preds <- get_cv_preds("excitement_index ~ score_difference + margin_1sthalf")
```

```{r five-fold, eval = FALSE}
bind_rows(mutate(ourvars_minusmargingames_cv_preds, type = "All minus Margin Games"),
          mutate(ourvars_cv_preds, type = "All"),
          mutate(all_int_cv_preds, type = "Some Interactions"),
          mutate(margin_only_cv_preds, type = "Score Differentials Only")) %>%
  group_by(type) %>%
  summarize(rmse = sqrt(mean((test_actual - test_preds)^2))) %>%
  mutate(type = fct_reorder(type, rmse)) %>%
  ggplot(aes(x = type, y = rmse)) +
  geom_point() + coord_flip() + theme_bw()
```

```{r}
all_lm <- lm(excitement_index ~ score_difference * margin_1sthalf + comeback + overtime, data = cfb)
summary(all_lm)
```

```{r}
all_lm2 <- lm(excitement_index ~ comeback + score_difference * margin_1sthalf, data = cfb)
summary(all_lm2)
```

```{r first-line-plot, eval = FALSE}
cfb %>%
  ggplot(aes(x = score_difference, y = excitement_index, color = comeback, shape = overtime, size = overtime)) + #maybe too much going on here
  geom_point() +
  #facet_wrap(~ week)
  theme_bw() 
```

Games that were comebacks/went to overtime/were a close game in terms of points were generally more exciting.


all_lm <- lm(excitement_index ~ score_difference * margin_1sthalf + comeback + overtime, data = cfb)
summary(all_lm)

```{r}
cfb %>%
  mutate(score_preds = predict(all_lm2)) %>%
  ggplot(aes(x = score_difference, 
             color = comeback)) +
  geom_point(aes(y = score_preds),
             alpha = 0.5) +
  theme_bw() +
  labs(x = "Score Difference", 
       y = "Predicted Excitement Index",
       title = "Comeback Games and Close Games are Predicted to be More Exciting") +
  geom_smooth(aes(y = excitement_index),
              method = "lm")
  ggsave("pred_ei_plot_ot.jpeg", width = 8, height = 6)
```

```{r}
all_lm_pred <- lm(excitement_index ~ score_difference * margin_1sthalf + comeback + overtime, data = cfb_filt)
```

```{r}
cfb_filt <- cfb %>%
  filter(margin_game >= 0 & margin_game <= 8) %>%
  mutate(score_preds = predict(all_lm_pred)) %>%
  ggplot(aes(x = score_difference, 
             color = overtime)) +
  geom_point(aes(y = score_preds),
             alpha = 0.75) +
  theme_bw() +
  labs(x = "Score Difference", 
       y = "Predicted Excitement Index",
       title = "Overtime Games and Close Games are Predicted to be More Exciting",
       caption = "Only games where score difference was within one possession") +
  geom_smooth(aes(y = excitement_index), method = "glm", level = .6)
  ggsave("pred_ei_plot_ot3.jpeg", width = 8, height = 6)
cfb_filt
```

Model diagnostics:

```{r}
cfb %>%
  mutate(init_preds = predict(all_lm2)) %>%
  ggplot(aes(x = excitement_index, y = init_preds)) +
  geom_point(alpha = 0.75) +
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", color = "red") +
  annotate(geom = "text", x = 8, y = 10, label="R-squared: 0.6528") +
  theme_bw() +
  labs(x = "Observed Excitement Index", y = "Predicted Excitement Index", title = "Predicted vs. Observed Excitement Index")
ggsave("pred_vs_obs.jpeg", width = 8, height = 6)
```

```{r}
#all_lm <- lm(excitement_index ~ score_difference * margin_1sthalf + comeback + overtime, data = cfb)
#plot(all_lm)

library(ggfortify)
testresid <- autoplot(all_lm) +
  theme_bw()
ggsave("testresid.jpeg", plot = testresid, width = 8, height = 6)
```

```{r}
ggcoef(all_lm, 
       vline = TRUE,
       vline_color = "red") + 
  theme_bw()
```

```{r}
vif(all_lm)
```
